---
title: "class-06"
format: pdf
editor: visual
date: 01/26/24
author: Iulia Rusu
---


all functions have 3 items:

- a name
- input arguements ( none, one or more )
- a body

a function to add two numbers
```{r}
sillyadd <- function(x) {
  x + 1
}
```

try out function
```{r}
sillyadd(10)
```

```{r}
sillyadd <- function (x, y= 1) { #If no y given do not specify y, it will be 1
 x + y
}

sillyadd(10)
sillyadd(10 + 3)

```

```{r}
#Example input vectors to start with 
student1<-c(100,100,100,100,100,100,100,90) 
student2<-c(100,NA,90,90,90,90,97,80) 
student3<-c(90,NA,NA,NA,NA,NA,NA,NA)
```

```{r}
```


```{r}
mean(student1)
min(student1)


```

```{r}
# Find lowest value
x <- student1
lowest_index <- which.min(x)
#exclude lowest value and find mean
mean(x[-lowest_index])

```
```{r}
# Find lowest value
student2
x <- student2
lowest_index <- which.min(x)
lowest_index
#exclude lowest value and find mean
mean(x[-lowest_index], na.rm=T)
```

```{r}
# Find lowest value
student3
x <- student3
x[is.na(x)] <- 0
lowest_index <- which.min(x)
lowest_index

#exclude lowest value and find mean
mean(x[-lowest_index], na.rm=T)
```


```{r}
grade <- function(x) { 
  x[is.na(x)] <- 0
  lowest_index <- which.min(x)
  lowest_index

#exclude lowest value and find mean
  mean(x[-lowest_index], na.rm=T)
  
}
```

```{r}
grade(student1)
```
Read class gradebook csv file from here: “https://tinyurl.com/gradeinput”
```{r}
url <- "https://tinyurl.com/gradeinput"

gradebook <- read.csv(url, row.names = 1)
```

```{r}
head(gradebook)
```
We can apply our new 'grade' function over either the rows or the columns of the gradebook. with MARGIN=1, or MARGIN=2
```{r}
apply(gradebook,1, grade)
```
Q2.Usingyourgrade()functionandthesuppliedgradebook,Whoisthetopscoringstudent overall inthegradebook? [3pts]
```{r}
averages <- apply(gradebook,1, grade)

which.max (averages)
```
Q3.Fromyouranalysisofthegradebook,whichhomeworkwastoughestonstudents(i.e.obtained thelowestscoresoverall? [2pts]

HW3

```{r}
apply(gradebook, 2, mean, na.rm=TRUE)
```
```{r}
grade <- function(x, drop.lowest= TRUE) { 
  x[is.na(x)] <- 0
  
  if(drop.lowest){
    ans <- mean(x[-which.min(x)])
  }
  else {
    
    ans <- mean(x)
  }
    ans

}


```

Q4. OptionalExtension: Fromyouranalysisof thegradebook,whichhomeworkwasmost predictiveofoverallscore(i.e.highestcorrelationwithaveragegradescore)? [1pt]
```{r}
gradebook$hw5
averages
mask <- gradebook
mask[is.na(mask)] <- 0
cor( mask$hw5, averages)
cor( mask$hw3, averages)

apply(mask, 2, cor, y =averages)
```

